---
title: "Comparison of rDNA references for realignment on MZ Twins"
---
  
```{r}
library(tidyverse)
source('../setup.R')
```


## Auxiliary functions

```{r}
adjust_pos <- function(pos, loop_size=2120) {
  if_else(pos<=loop_size, pos-loop_size-1, pos-loop_size)
}
```

```{r}
read_lofreq <- function(fd, fn='snps', loop_size=2120) {
  bind_rows(
    lapply(
      X=list.files(fd),
      FUN=function(x) { 
        if(!file.exists(
          file.path(fd, x, paste0(x, ".", fn, ".vcf")))) return(NULL)
        read_tsv(
          file.path(fd, x, paste0(x, ".", fn, ".vcf")), 
          comment = "#",
          col_names=F,
          col_types=cols()) %>%
          mutate(sample=x)})) %>%
    select(sample, pos=X2, ref=X4, alt=X5, info=X8) %>%
    separate(info, into=c('dp', 'af'), sep=';', extra='drop') %>%
    mutate(af=as.numeric(gsub('AF=', '', af)),
           dp=as.numeric(gsub('DP=', '', dp))) %>%
    mutate(pos_ref=adjust_pos(pos, loop_size=loop_size))
}
```

```{r}
compare_twins_af_var <- function(df, the_var = variants[1]) {
  df_aux <- df |> filter(var == the_var)
  if(nrow(df_aux) < 3) return(NULL)
  cor.test(df_aux$af_1, df_aux$af_2) |>
    broom::tidy() |>
    mutate(
      Variant = the_var,
      N = nrow(df_aux)) |>
    relocate(Variant, N)}
```

## Relative pairs

```{r}
relatives <- read_csv(file.path(root_fd, '..', 'data', 'pheno', 'ukb_rel_dist.csv'), 
                      col_types = cols())
```

```{r}
twins <- relatives |> 
  filter(grepl('MZ', dist)) |>
  select(relation, eid_1 = ID1, eid_2 = ID2)
```

```{r}
ukb_demo_wb <- read_csv(file.path(root_fd, '..', 'data', 'pheno', 'ukb_demo.csv'),
                     col_types = cols()) |>
  filter(Eth_Back == 1001)
```

```{r}
twins_wb <- twins |> filter(
  eid_1 %in% ukb_demo_wb$eid,
  eid_2 %in% ukb_demo_wb$eid)
```

```{r}
twins_wb_eids <- c(twins_wb$eid_1, twins_wb$eid_2)
```


## Full realignment to our tailored Hg38 + rDNA

### First Sequencing Release

### deCODE 1 + Sanger 1

```{r}
ifd_s1_s1 <- file.path(
  root_fd, '01_validation', 'data', '01_realigned_variants',
  '01_bt2_hg38' , 'twins', 's1_s1')
ifd_d1_d1 <- file.path(
  root_fd, '01_validation', 'data', '01_realigned_variants',
  '01_bt2_hg38' , 'twins', 'd1_d1')
```

```{r}
twins_s1_s1_many_vars <- read_lofreq(ifd_s1_s1, 'many')
twins_d1_d1_many_vars <- read_lofreq(ifd_d1_d1, 'many')
```

```{r}
twins_s1_s1_unfilt_vars <- read_lofreq(ifd_s1_s1, 'unfilt')
twins_d1_d1_unfilt_vars <- read_lofreq(ifd_d1_d1, 'unfilt')
```

```{r}
twins_s1_s1_hc_vars <- read_lofreq(ifd_s1_s1, 'hc')
twins_d1_d1_hc_vars <- read_lofreq(ifd_d1_d1, 'hc')
```


```{r}
twins_ds1_ds1_unfilt_vars <- bind_rows(
  twins_d1_d1_unfilt_vars |> mutate(SC = 'deCODE'),
  twins_s1_s1_unfilt_vars |> mutate(SC = 'Sanger'))
```

```{r}
twins_ds1_ds1_hc_vars <- bind_rows(
  twins_d1_d1_hc_vars |> mutate(SC = 'deCODE'),
  twins_s1_s1_hc_vars |> mutate(SC = 'Sanger'))
```

```{r}
twins_ds1_ds1_hc_vars_distinct <- twins_ds1_ds1_hc_vars |>
  select(pos_ref, ref, alt) |> distinct() |>
  arrange(pos_ref, ref, alt) |>
  mutate(var = paste(pos_ref, ref, alt, sep = '_'))
```

```{r}
twins_ds1_ds1_afs <- twins_ds1_ds1_unfilt_vars |> 
  inner_join(twins_ds1_ds1_hc_vars_distinct,
             by = c('pos_ref', 'ref', 'alt')) |>
  mutate(eid = as.numeric(sample)) |>
  select(-c(sample, pos)) |>
  relocate(eid)
```

```{r}
twins_ds1_ds1_hc_vars_list <- twins_ds1_ds1_hc_vars_distinct |> pull(var)
```

### Second Sequencing Release

```{r}
ifd_s2_s2 <- file.path(
  root_fd, '01_validation', 'data', '01_realigned_variants',
  '01_bt2_hg38' , 'twins', 's2_s2')
ifd_d2_d2 <- file.path(
  root_fd, '01_validation', 'data', '01_realigned_variants',
  '01_bt2_hg38' , 'twins', 'd2_d2')
```

```{r}
twins_d2_d2_many_vars <- read_lofreq(ifd_d2_d2, 'many')
twins_s2_s2_many_vars <- read_lofreq(ifd_s2_s2, 'many')
```

```{r}
twins_d2_d2_unfilt_vars <- read_lofreq(ifd_d2_d2, 'unfilt')
twins_s2_s2_unfilt_vars <- read_lofreq(ifd_s2_s2, 'unfilt')
```

```{r}
twins_d2_d2_hc_vars <- read_lofreq(ifd_d2_d2, 'hc')
twins_s2_s2_hc_vars <- read_lofreq(ifd_s2_s2, 'hc')
```

```{r}
twins_ds2_ds2_unfilt_vars <- bind_rows(
  twins_d2_d2_unfilt_vars |> mutate(SC = 'deCODE'),
  twins_s2_s2_unfilt_vars |> mutate(SC = 'Sanger'))
```

```{r}
twins_ds2_ds2_hc_vars <- bind_rows(
  twins_d2_d2_hc_vars |> mutate(SC = 'deCODE'),
  twins_s2_s2_hc_vars |> mutate(SC = 'Sanger'))
```

```{r}
twins_ds2_ds2_hc_vars_distinct <- twins_ds2_ds2_hc_vars |>
  select(pos_ref, ref, alt) |> distinct() |>
  arrange(pos_ref, ref, alt) |>
  mutate(var = paste(pos_ref, ref, alt, sep = '_'))
```

```{r}
twins_ds2_ds2_afs <- twins_ds2_ds2_unfilt_vars |> 
  inner_join(twins_ds2_ds2_hc_vars_distinct,
             by = c('pos_ref', 'ref', 'alt')) |>
  mutate(eid = as.numeric(sample)) |>
  select(-c(sample, pos)) |>
  relocate(eid)
```

```{r}
twins_ds2_ds2_hc_vars_list <- twins_ds2_ds2_hc_vars_distinct |> pull(var)
```


#### Combine First and Second release

```{r}
twins_hc_vars_distinct <- bind_rows(
  twins_ds1_ds1_hc_vars_distinct,
  twins_ds2_ds2_hc_vars_distinct) |>
  arrange(pos_ref, ref, alt) |>
  distinct()
```

```{r}
twins_afs <- bind_rows(
  twins_ds1_ds1_unfilt_vars,
  twins_ds2_ds2_unfilt_vars) |> 
  inner_join(twins_hc_vars_distinct,
             by = c('pos_ref', 'ref', 'alt')) |>
  mutate(eid = as.numeric(sample)) |>
  select(-c(sample, pos)) |>
  relocate(eid) |>
  arrange(pos_ref, ref, alt)
```

```{r}
twins_hc_vars_list <- twins_hc_vars_distinct |> pull(var)
```

```{r}
summary_01_af <-  twins_afs |> group_by(pos_ref, ref, alt) |>
      summarise(dp_mean = mean(dp), dp_median = median(dp),
                dp_low = min(dp), dp_high = max(dp),
                af_mean = mean(dp), af_median = median(af),
                af_low = min(af), af_high = max(af),
                .groups = 'drop')
```

## Full realignment to looped KY962518.1

### First Release


```{r}
ifd_s1_s1_05 <- file.path(
  root_fd, '01_validation', 'data', '01_realigned_variants',
  '05_bt2_ky' , 'twins', 's1_s1')
ifd_d1_d1_05 <- file.path(
  root_fd, '01_validation', 'data', '01_realigned_variants',
  '05_bt2_ky' , 'twins', 'd1_d1')
```


```{r}
twins_d1_d1_05_many_vars <- read_lofreq(ifd_d1_d1_05, 'many')
twins_s1_s1_05_many_vars <- read_lofreq(ifd_s1_s1_05, 'many')
```

```{r}
twins_d1_d1_05_unfilt_vars <- read_lofreq(
  ifd_d1_d1_05, 'unfilt')
twins_s1_s1_05_unfilt_vars <- read_lofreq(
  ifd_s1_s1_05, 'unfilt')
```

```{r}
twins_d1_d1_05_hc_vars <- read_lofreq(
  ifd_d1_d1_05, 'hc')
twins_s1_s1_05_hc_vars <- read_lofreq(
  ifd_s1_s1_05, 'hc')
```

```{r}
twins_ds1_ds1_05_unfilt_vars <- bind_rows(
  twins_d1_d1_05_unfilt_vars |> mutate(SC = 'deCODE'),
  twins_s1_s1_05_unfilt_vars |> mutate(SC = 'Sanger'))
```

```{r}
twins_ds1_ds1_05_hc_vars <- bind_rows(
  twins_d1_d1_05_hc_vars |> mutate(SC = 'deCODE'),
  twins_s1_s1_05_hc_vars |> mutate(SC = 'Sanger'))
```

```{r}
twins_ds1_ds1_05_hc_vars_distinct <- twins_ds1_ds1_05_hc_vars |>
  select(pos_ref, ref, alt) |> distinct() |>
  arrange(pos_ref, ref, alt) |>
  mutate(var = paste(pos_ref, ref, alt, sep = '_'))
```

```{r}
twins_ds1_ds1_05_afs <- twins_ds1_ds1_05_unfilt_vars |> 
  inner_join(twins_ds1_ds1_05_hc_vars_distinct,
             by = c('pos_ref', 'ref', 'alt')) |>
  mutate(eid = as.numeric(sample)) |>
  select(-c(sample, pos)) |>
  relocate(eid)
```

```{r}
twins_ds1_ds1_05_hc_vars_list <- twins_ds1_ds1_05_hc_vars_distinct |> pull(var)
```

### Second Release

```{r}
ifd_s2_s2_05 <- file.path(
  root_fd, '01_validation', 'data', '01_realigned_variants',
  '05_bt2_ky' , 'twins', 's2_s2')
ifd_d2_d2_05 <- file.path(
  root_fd, '01_validation', 'data', '01_realigned_variants',
  '05_bt2_ky' , 'twins', 'd2_d2')
```

```{r}
twins_d2_d2_05_many_vars <- read_lofreq(ifd_d2_d2_05, 'many')
twins_s2_s2_05_many_vars <- read_lofreq(ifd_s2_s2_05, 'many')
```

```{r}
twins_d2_d2_05_unfilt_vars <- read_lofreq(
  ifd_d2_d2_05, 'unfilt')
twins_s2_s2_05_unfilt_vars <- read_lofreq(
  ifd_s2_s2_05, 'unfilt')
```

```{r}
twins_d2_d2_05_hc_vars <- read_lofreq(
  ifd_d2_d2_05, 'hc')
twins_s2_s2_05_hc_vars <- read_lofreq(
  ifd_s2_s2_05, 'hc')
```

```{r}
twins_ds2_ds2_05_unfilt_vars <- bind_rows(
  twins_d2_d2_05_unfilt_vars |> mutate(SC = 'deCODE'),
  twins_s2_s2_05_unfilt_vars |> mutate(SC = 'Sanger'))
```

```{r}
twins_ds2_ds2_05_hc_vars <- bind_rows(
  twins_d2_d2_05_hc_vars |> mutate(SC = 'deCODE'),
  twins_s2_s2_05_hc_vars |> mutate(SC = 'Sanger'))
```

```{r}
twins_ds2_ds2_05_hc_vars_distinct <- twins_ds2_ds2_05_hc_vars |>
  select(pos_ref, ref, alt) |> distinct() |>
  arrange(pos_ref, ref, alt) |>
  mutate(var = paste(pos_ref, ref, alt, sep = '_'))
```

```{r}
twins_ds2_ds2_05_afs <- twins_ds2_ds2_05_unfilt_vars |> 
  inner_join(twins_ds2_ds2_05_hc_vars_distinct,
             by = c('pos_ref', 'ref', 'alt')) |>
  mutate(eid = as.numeric(sample)) |>
  select(-c(sample, pos)) |>
  relocate(eid)
```

```{r}
twins_ds2_ds2_05_hc_vars_list <- twins_ds2_ds2_05_hc_vars_distinct |> pull(var)
```

### Combine First and Second release

```{r}
twins_05_hc_vars_distinct <- bind_rows(
  twins_ds1_ds1_05_hc_vars_distinct,
  twins_ds2_ds2_05_hc_vars_distinct) |>
  arrange(pos_ref, ref, alt) |>
  distinct()
```

```{r}
twins_05_afs <- bind_rows(
  twins_ds1_ds1_05_unfilt_vars,
  twins_ds2_ds2_05_unfilt_vars) |> 
  inner_join(twins_05_hc_vars_distinct,
             by = c('pos_ref', 'ref', 'alt')) |>
  mutate(eid = as.numeric(sample)) |>
  select(-c(sample, pos)) |>
  relocate(eid) |>
  arrange(pos_ref, ref, alt)
```

```{r}
twins_05_hc_vars_list <- twins_hc_vars_distinct |> pull(var)
```


#### Summary

```{r}
summary_05_af <-  twins_05_afs |> group_by(pos_ref, ref, alt) |>
      summarise(dp_mean = mean(dp), dp_median = median(dp),
                dp_low = min(dp), dp_high = max(dp),
                af_mean = mean(dp), af_median = median(af),
                af_low = min(af), af_high = max(af),
                .groups = 'drop')
```

## Comparison of approaches

### 01 vs 05

```{r}
summary_01_05_af <- full_join(
  summary_01_af,
  summary_05_af,
  by = c('pos_ref', 'ref', 'alt'), suffix = c('_01', '_05')) |>
  arrange(pos_ref, ref, alt) |>
  select(pos_ref, ref, alt, 
         starts_with('af_median'))
```

```{r}
summary_01_05_af |>
  mutate(imputed = (is.na(af_median_01) | is.na(af_median_05))) |>
  replace_na(list(af_median_01 = 0, af_median_05 = 0)) |>
  ggplot(aes(x = af_median_01, y = af_median_05)) +
  geom_point(aes(fill = imputed), alpha = 0.5, shape = 21) +
  ggpubr::stat_cor(
    aes(label = cor_format(after_stat(r), after_stat(p)))) +
  scale_fill_manual(values = c('FALSE' = 'grey40', 'TRUE' = 'firebrick')) +
  theme_bw() +
  labs(x = "Median Variant Frequency in tailored Hg38+rDNA",
       y = "Median Variant Frequency in rDNA only") +
  theme_pub() +
  theme(
    aspect.ratio = 1,
    legend.position = 'none',
    panel.grid = element_blank(),
    panel.border = element_rect(linewidth = 1.1),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 13))
```


## Rescue of frequencies

```{r}
vars_05 <- twins_05_hc_vars_distinct |>
  select(pos_ref, ref, alt) |>
  distinct()
```

```{r}
afs_many_05 <- bind_rows(
  twins_d1_d1_05_many_vars |> mutate(release = 'd1'),
  twins_s1_s1_05_many_vars |> mutate(release = 's1'),
  twins_d2_d2_05_many_vars |> mutate(release = 'd2'),
  twins_s2_s2_05_many_vars |> mutate(release = 's2')) |>
    inner_join(vars_05, by = c('pos_ref', 'ref', 'alt')) |>
  mutate(eid = as.numeric(sample)) |>
  select(-sample) |>
  mutate(var = paste(pos_ref, ref, alt, sep = '_'))
```

```{r}
twins_05_afs_many <- twins |>
  inner_join(afs_many_05,
             by = c('eid_1' = 'eid')) |>
  inner_join(afs_many_05,
             by = c('eid_2' = 'eid', 'pos_ref', 'ref', 'alt', 'var'),
             suffix = c('_1', '_2')) |>
  arrange(pos_ref, ref, alt) 
```

```{r}
twins_05_afs_many |>
    ggplot(aes(x = af_1, y = af_2)) +
    geom_point(alpha = 0.5) +
    ggpubr::stat_cor(size = 2) +
    labs(x = 'Allele Frequency in Twin 1',
         y = 'Allele Frequency in Twin 2') +
  facet_grid(cols = vars(release_1)) +
    theme_bw() +
    theme(panel.grid = element_blank(),
          aspect.ratio = 1)
```


## Full realignment to Paralkar's Lab Hg38 + rDNA

### deCODE 2 vs deCODE 2


```{r}
ifd_d2_d2_03 <- file.path(
  root_fd, '01_validation', 'data', '01_realigned_variants',
  '05_bt2_hg38_paralkar' , 'twins', 'd2_d2')
```


```{r}
twins_d2_d2_03_many_vars <- read_lofreq(ifd_d2_d2_03, 'many', 
                                        loop_size = (44838 - 35500))
```

```{r}
twins_d2_d2_03_unfilt_vars <- read_lofreq(
  ifd_d2_d2_03, 'unfilt', loop_size = (44838 - 35500))
```

```{r}
twins_d2_d2_03_hc_vars <- read_lofreq(
  ifd_d2_d2_03, 'hc', loop_size = (44838 - 35500))
```

```{r}
twins_d2_d2_03_hc_vars_distinct <- twins_d2_d2_03_hc_vars |>
  select(pos_ref, ref, alt) |> distinct() |>
  arrange(pos_ref, ref, alt) |>
  mutate(var = paste(pos_ref, ref, alt, sep = '_'))
```

```{r}
twins_d2_d2_03_afs <- twins_d2_d2_03_unfilt_vars |> 
  inner_join(twins_d2_d2_03_hc_vars_distinct,
             by = c('pos_ref', 'ref', 'alt')) |>
  mutate(eid = as.numeric(sample)) |>
  select(-c(sample, pos)) |>
  relocate(eid)
```

```{r}
twins_d2_d2_03_hc_vars_list <- twins_d2_d2_03_hc_vars_distinct |> pull(var)
```

```{r}
twins_d2_d2_03_real_afs_comp <- twins |>
  inner_join(twins_d2_d2_03_afs,
             by = c('eid_1' = 'eid')) |>
  inner_join(twins_d2_d2_03_afs,
             by = c('eid_2' = 'eid', 'pos_ref', 'ref', 'alt', 'var'),
             suffix = c('_1', '_2')) |>
  arrange(pos_ref, ref, alt) |>
  mutate(var = forcats::fct_inorder(var))
```

```{r}
twins_d2_d2_03_real_afs_comp |>
    rename(variant = var) |>
    group_by(variant) |>
    filter(pos_ref == -424) |>
    ungroup() |>
    ggplot(aes(x = af_1, y = af_2)) +
    geom_point(alpha = 0.5) +
    ggpubr::stat_cor(size = 2) +
    facet_wrap(. ~ variant, scales = 'free') +
    labs(x = 'Allele Frequency in Twin 1',
         y = 'Allele Frequency in Twin 2') +
    theme_bw() +
    theme(panel.grid = element_blank(),
          aspect.ratio = 1)
```


```{r}
twins_d2_d2_03_real_afs_cor <- bind_rows(
  lapply(X = twins_d2_d2_03_hc_vars_list,
         FUN = function(x) {
           compare_twins_af_var(df = twins_d2_d2_03_real_afs_comp,
                                the_var = x) }))
```



### Sanger 2 vs Sanger 2

```{r}
ifd_s2_s2_03 <- file.path(
  root_fd, '01_validation', 'data', '01_realigned_variants',
  '05_bt2_hg38_paralkar' , 'twins', 's2_s2')
```

```{r}
twins_s2_s2_03_many_vars <- read_lofreq(ifd_s2_s2_03, 'many',
                                        loop_size =  (44838 - 35500))
```

```{r}
twins_s2_s2_03_unfilt_vars <- read_lofreq(
  ifd_s2_s2_03, 'unfilt', loop_size = (44838 - 35500))
```

```{r}
twins_s2_s2_03_hc_vars <- read_lofreq(
  ifd_s2_s2_03, 'hc', loop_size = (44838 - 35500))
```

```{r}
twins_s2_s2_03_hc_vars_distinct <- twins_s2_s2_03_hc_vars |>
  select(pos_ref, ref, alt) |> distinct() |>
  arrange(pos_ref, ref, alt) |>
  mutate(var = paste(pos_ref, ref, alt, sep = '_'))
```

```{r}
twins_s2_s2_03_afs <- twins_s2_s2_03_unfilt_vars |> 
  inner_join(twins_s2_s2_03_hc_vars_distinct,
             by = c('pos_ref', 'ref', 'alt')) |>
  mutate(eid = as.numeric(sample)) |>
  select(-c(sample, pos)) |>
  relocate(eid)
```

```{r}
twins_s2_s2_03_hc_vars_list <- twins_s2_s2_03_hc_vars_distinct |> pull(var)
```

```{r}
twins_s2_s2_03_real_afs_comp <- twins |>
  inner_join(twins_s2_s2_03_afs,
             by = c('eid_1' = 'eid')) |>
  inner_join(twins_s2_s2_03_afs,
             by = c('eid_2' = 'eid', 'pos_ref', 'ref', 'alt', 'var'),
             suffix = c('_1', '_2')) |>
  arrange(pos_ref, ref, alt) |>
  mutate(var = forcats::fct_inorder(var))
```

```{r}
twins_s2_s2_03_real_afs_comp |>
    rename(variant = var) |>
    group_by(variant) |>
    filter(pos_ref == -424) |>
    ungroup() |>
    ggplot(aes(x = af_1, y = af_2)) +
    geom_point(alpha = 0.5) +
    ggpubr::stat_cor(size = 2) +
    facet_wrap(. ~ variant, scales = 'free') +
    labs(x = 'Allele Frequency in Twin 1',
         y = 'Allele Frequency in Twin 2') +
    theme_bw() +
    theme(panel.grid = element_blank(),
          aspect.ratio = 1)
```


```{r}
twins_s2_s2_03_real_afs_cor <- bind_rows(
  lapply(X = twins_s2_s2_03_hc_vars_list,
         FUN = function(x) {
           compare_twins_af_var(df = twins_s2_s2_03_real_afs_comp,
                                the_var = x) }))
```


### Combined deCODE 2 / Sanger 2

```{r}
twins_ds2_ds2_03_unfilt_vars <- bind_rows(
  twins_d2_d2_03_unfilt_vars |> mutate(SC = 'deCODE'),
  twins_s2_s2_03_unfilt_vars |> mutate(SC = 'Sanger'))
```

```{r}
twins_ds2_ds2_03_hc_vars <- bind_rows(
  twins_d2_d2_03_hc_vars |> mutate(SC = 'deCODE'),
  twins_s2_s2_03_hc_vars |> mutate(SC = 'Sanger'))
```

```{r}
twins_ds2_ds2_03_hc_vars_distinct <- twins_ds2_ds2_03_hc_vars |>
  select(pos_ref, ref, alt) |> distinct() |>
  arrange(pos_ref, ref, alt) |>
  mutate(var = paste(pos_ref, ref, alt, sep = '_'))
```

```{r}
twins_ds2_ds2_03_afs <- twins_ds2_ds2_03_unfilt_vars |> 
  inner_join(twins_ds2_ds2_03_hc_vars_distinct,
             by = c('pos_ref', 'ref', 'alt')) |>
  mutate(eid = as.numeric(sample)) |>
  select(-c(sample, pos)) |>
  relocate(eid)
```

```{r}
twins_ds2_ds2_03_hc_vars_list <- twins_ds2_ds2_03_hc_vars_distinct |> pull(var)
```

```{r}
twins_ds2_ds2_03_real_afs_comp <- twins |>
  inner_join(twins_ds2_ds2_03_afs,
             by = c('eid_1' = 'eid')) |>
  inner_join(twins_ds2_ds2_03_afs,
             by = c('eid_2' = 'eid', 'pos_ref', 'ref', 'alt', 'var'),
             suffix = c('_1', '_2')) |>
  arrange(pos_ref, ref, alt) |>
  mutate(var = forcats::fct_inorder(var))
```

```{r}
twins_ds2_ds2_03_real_afs_comp |>
    rename(variant = var) |>
    group_by(variant) |>
    filter(pos_ref == -424) |>
    ungroup() |>
    ggplot(aes(x = af_1, y = af_2)) +
    geom_point(alpha = 0.5) +
    ggpubr::stat_cor(size = 2) +
    facet_wrap(. ~ variant, scales = 'free') +
    labs(x = 'Allele Frequency in Twin 1',
         y = 'Allele Frequency in Twin 2') +
    theme_bw() +
    theme(panel.grid = element_blank(),
          aspect.ratio = 1)
```

```{r}
twins_ds2_ds2_03_real_afs_cor <- bind_rows(
  lapply(X = twins_ds2_ds2_03_hc_vars_list,
         FUN = function(x) {
           compare_twins_af_var(df = twins_ds2_ds2_03_real_afs_comp,
                                the_var = x) }))
```

#### Summary

```{r}
summary_03 <- twins_ds2_ds2_03_real_afs_comp |>
  group_by(pos_ref, ref, alt) |>
  summarise(num_pairs = n_distinct(relation),
            .groups = 'drop') |>
  left_join(
    twins_ds2_ds2_03_real_afs_cor |>
      select(Variant, R = estimate) |>
      separate(Variant, into = c('pos_ref', 'ref', 'alt'), sep = '_') |>
      mutate(pos_ref = as.numeric(pos_ref)),
    by = c('pos_ref', 'alt', 'ref')) |>
  inner_join(
    twins_ds2_ds2_03_afs |> group_by(pos_ref, ref, alt) |>
      summarise(dp_mean = mean(dp), dp_median = median(dp),
                dp_low = min(dp), dp_high = max(dp),
                af_mean = mean(dp), af_median = median(af),
                af_low = min(af), af_high = max(af),
                .groups = 'drop'),
    by = c('pos_ref', 'ref', 'alt'))
```


## Full realignment to our Paralkar's Lab T2T + rDNA

### deCODE 2 vs deCODE 2

```{r}
ifd_d2_d2_04 <- file.path(
  root_fd, '01_validation', 'data', '01_realigned_variants',
  '04_bt2_t2t_paralkar' , 'twins', 'd2_d2')
```

```{r}
twins_d2_d2_04_many_vars <- read_lofreq(ifd_d2_d2_03, 'many',
                                        loop_size = (44838 - 35500))
```

```{r}
twins_d2_d2_04_unfilt_vars <- read_lofreq(
  ifd_d2_d2_04, 'unfilt', loop_size = (44838 - 35500))
```

```{r}
twins_d2_d2_04_hc_vars <- read_lofreq(
  ifd_d2_d2_04, 'hc', loop_size = (44838 - 35500))
```

```{r}
twins_d2_d2_04_hc_vars_distinct <- twins_d2_d2_04_hc_vars |>
  select(pos_ref, ref, alt) |> distinct() |>
  arrange(pos_ref, ref, alt) |>
  mutate(var = paste(pos_ref, ref, alt, sep = '_'))
```

```{r}
twins_d2_d2_04_afs <- twins_d2_d2_04_unfilt_vars |> 
  inner_join(twins_d2_d2_04_hc_vars_distinct,
             by = c('pos_ref', 'ref', 'alt')) |>
  mutate(eid = as.numeric(sample)) |>
  select(-c(sample, pos)) |>
  relocate(eid)
```

```{r}
twins_d2_d2_04_hc_vars_list <- twins_d2_d2_04_hc_vars_distinct |> pull(var)
```

```{r}
twins_d2_d2_04_real_afs_comp <- twins |>
  inner_join(twins_d2_d2_04_afs,
             by = c('eid_1' = 'eid')) |>
  inner_join(twins_d2_d2_04_afs,
             by = c('eid_2' = 'eid', 'pos_ref', 'ref', 'alt', 'var'),
             suffix = c('_1', '_2')) |>
  arrange(pos_ref, ref, alt) |>
  mutate(var = forcats::fct_inorder(var))
```

```{r}
twins_d2_d2_04_real_afs_comp |>
    rename(variant = var) |>
    group_by(variant) |>
    filter(pos_ref == -424) |>
    ungroup() |>
    ggplot(aes(x = af_1, y = af_2)) +
    geom_point(alpha = 0.5) +
    ggpubr::stat_cor(size = 2) +
    facet_wrap(. ~ variant, scales = 'free') +
    labs(x = 'Allele Frequency in Twin 1',
         y = 'Allele Frequency in Twin 2') +
    theme_bw() +
    theme(panel.grid = element_blank(),
          aspect.ratio = 1)
```


```{r}
twins_d2_d2_04_real_afs_cor <- bind_rows(
  lapply(X = twins_d2_d2_04_hc_vars_list,
         FUN = function(x) {
           compare_twins_af_var(df = twins_d2_d2_04_real_afs_comp,
                                the_var = x) }))
```


### Sanger 2 vs Sanger 2

```{r}
ifd_s2_s2_04 <- file.path(
  root_fd, '01_validation', 'data', '01_realigned_variants',
  '04_bt2_t2t_paralkar' , 'twins', 's2_s2')
```

```{r}
twins_s2_s2_04_many_vars <- read_lofreq(ifd_s2_s2_04, 'many',
                                        loop_size = (44838 - 35500) )
```

```{r}
twins_s2_s2_04_unfilt_vars <- read_lofreq(
  ifd_s2_s2_04, 'unfilt', loop_size = (44838 - 35500))
```

```{r}
twins_s2_s2_04_hc_vars <- read_lofreq(
  ifd_s2_s2_04, 'hc', loop_size = (44838 - 35500))
```

```{r}
twins_s2_s2_04_hc_vars_distinct <- twins_s2_s2_04_hc_vars |>
  select(pos_ref, ref, alt) |> distinct() |>
  arrange(pos_ref, ref, alt) |>
  mutate(var = paste(pos_ref, ref, alt, sep = '_'))
```

```{r}
twins_s2_s2_04_afs <- twins_s2_s2_04_unfilt_vars |> 
  inner_join(twins_s2_s2_04_hc_vars_distinct,
             by = c('pos_ref', 'ref', 'alt')) |>
  mutate(eid = as.numeric(sample)) |>
  select(-c(sample, pos)) |>
  relocate(eid)
```

```{r}
twins_s2_s2_04_hc_vars_list <- twins_s2_s2_04_hc_vars_distinct |> pull(var)
```

```{r}
twins_s2_s2_04_real_afs_comp <- twins |>
  inner_join(twins_s2_s2_04_afs,
             by = c('eid_1' = 'eid')) |>
  inner_join(twins_s2_s2_04_afs,
             by = c('eid_2' = 'eid', 'pos_ref', 'ref', 'alt', 'var'),
             suffix = c('_1', '_2')) |>
  arrange(pos_ref, ref, alt) |>
  mutate(var = forcats::fct_inorder(var))
```

```{r}
twins_s2_s2_04_real_afs_comp |>
    rename(variant = var) |>
    group_by(variant) |>
    filter(pos_ref == -424) |>
    ungroup() |>
    ggplot(aes(x = af_1, y = af_2)) +
    geom_point(alpha = 0.5) +
    ggpubr::stat_cor(size = 2) +
    facet_wrap(. ~ variant, scales = 'free') +
    labs(x = 'Allele Frequency in Twin 1',
         y = 'Allele Frequency in Twin 2') +
    theme_bw() +
    theme(panel.grid = element_blank(),
          aspect.ratio = 1)
```


```{r}
twins_s2_s2_04_real_afs_cor <- bind_rows(
  lapply(X = twins_s2_s2_04_hc_vars_list,
         FUN = function(x) {
           compare_twins_af_var(df = twins_s2_s2_04_real_afs_comp,
                                the_var = x) }))
```



### Combined deCODE 2 / Sanger 2

```{r}
twins_ds2_ds2_04_unfilt_vars <- bind_rows(
  twins_d2_d2_04_unfilt_vars |> mutate(SC = 'deCODE'),
  twins_s2_s2_04_unfilt_vars |> mutate(SC = 'Sanger'))
```

```{r}
twins_ds2_ds2_04_hc_vars <- bind_rows(
  twins_d2_d2_04_hc_vars |> mutate(SC = 'deCODE'),
  twins_s2_s2_04_hc_vars |> mutate(SC = 'Sanger'))
```

```{r}
twins_ds2_ds2_04_hc_vars_distinct <- twins_ds2_ds2_04_hc_vars |>
  select(pos_ref, ref, alt) |> distinct() |>
  arrange(pos_ref, ref, alt) |>
  mutate(var = paste(pos_ref, ref, alt, sep = '_'))
```

```{r}
twins_ds2_ds2_04_afs <- twins_ds2_ds2_04_unfilt_vars |> 
  inner_join(twins_ds2_ds2_04_hc_vars_distinct,
             by = c('pos_ref', 'ref', 'alt')) |>
  mutate(eid = as.numeric(sample)) |>
  select(-c(sample, pos)) |>
  relocate(eid)
```

```{r}
twins_ds2_ds2_04_hc_vars_list <- twins_ds2_ds2_04_hc_vars_distinct |> pull(var)
```

```{r}
twins_ds2_ds2_04_real_afs_comp <- twins |>
  inner_join(twins_ds2_ds2_04_afs,
             by = c('eid_1' = 'eid')) |>
  inner_join(twins_ds2_ds2_04_afs,
             by = c('eid_2' = 'eid', 'pos_ref', 'ref', 'alt', 'var'),
             suffix = c('_1', '_2')) |>
  arrange(pos_ref, ref, alt) |>
  mutate(var = forcats::fct_inorder(var))
```

```{r}
twins_ds2_ds2_04_real_afs_comp |>
    rename(variant = var) |>
    group_by(variant) |>
    filter(pos_ref == -424) |>
    ungroup() |>
    ggplot(aes(x = af_1, y = af_2)) +
    geom_point(alpha = 0.5) +
    ggpubr::stat_cor(size = 2) +
    facet_wrap(. ~ variant, scales = 'free') +
    labs(x = 'Allele Frequency in Twin 1',
         y = 'Allele Frequency in Twin 2') +
    theme_bw() +
    theme(panel.grid = element_blank(),
          aspect.ratio = 1)
```

```{r}
twins_ds2_ds2_04_real_afs_cor <- bind_rows(
  lapply(X = twins_ds2_ds2_04_hc_vars_list,
         FUN = function(x) {
           compare_twins_af_var(df = twins_ds2_ds2_04_real_afs_comp,
                                the_var = x) }))
```

#### Summary

```{r}
summary_04 <- twins_ds2_ds2_04_real_afs_comp |>
  group_by(pos_ref, ref, alt) |>
  summarise(num_pairs = n_distinct(relation),
            .groups = 'drop') |>
  left_join(
    twins_ds2_ds2_04_real_afs_cor |>
      select(Variant, R = estimate) |>
      separate(Variant, into = c('pos_ref', 'ref', 'alt'), sep = '_') |>
      mutate(pos_ref = as.numeric(pos_ref)),
    by = c('pos_ref', 'alt', 'ref')) |>
  inner_join(
    twins_ds2_ds2_04_afs |> group_by(pos_ref, ref, alt) |>
      summarise(dp_mean = mean(dp), dp_median = median(dp),
                dp_low = min(dp), dp_high = max(dp),
                af_mean = mean(dp), af_median = median(af),
                af_low = min(af), af_high = max(af),
                .groups = 'drop'),
    by = c('pos_ref', 'ref', 'alt'))
```


### 01 vs 03

```{r}
summary_01_03 <- full_join(
  summary_01,
  summary_03,
  by = c('pos_ref', 'ref', 'alt'), suffix = c('_01', '_03')) |>
  arrange(pos_ref, ref, alt) |>
  select(pos_ref, ref, alt, 
         starts_with('num_pairs'), 
         starts_with('R_'),
         starts_with('af_median'))
```


```{r}
summary_01_03 |>
  mutate(imputed = (is.na(R_01) | is.na(R_03))) |>
  replace_na(list(R_01 = 0, R_03 = 0)) |>
  ggplot(aes(x = R_01, y = R_03)) +
  geom_point(aes(fill = imputed), alpha = 0.5, shape = 21) +
  scale_fill_manual(values = c('FALSE' = 'grey40', 'TRUE' = 'firebrick')) +
  theme_bw() +
  labs(x = "Intra-twin Pearson's R in tailored Hg38+rDNA",
       y = "Intra-twin Pearson's R in Paralkar's Hg38+rDNA") +
  theme(
    aspect.ratio = 1,
    legend.position = 'none')
```

```{r}
summary_01_03 |>
  mutate(imputed = (is.na(af_median_01) | is.na(af_median_03))) |>
  replace_na(list(af_median_01 = 0, af_median_03 = 0)) |>
  ggplot(aes(x = af_median_01, y = af_median_03)) +
  geom_point(aes(fill = imputed), alpha = 0.5, shape = 21) +
  scale_fill_manual(values = c('FALSE' = 'grey40', 'TRUE' = 'firebrick')) +
  theme_bw() +
  labs(x = "Median Variant Frequency in tailored Hg38+rDNA",
       y = "Median Variant Frequency in Paralkar's Hg38+rDNA") +
  theme(
    aspect.ratio = 1,
    legend.position = 'none')
```

### 01 vs 04

```{r}
summary_01_04 <- full_join(
  summary_01,
  summary_04,
  by = c('pos_ref', 'ref', 'alt'), suffix = c('_01', '_04')) |>
  arrange(pos_ref, ref, alt) |>
  select(pos_ref, ref, alt, 
         starts_with('num_pairs'), 
         starts_with('R_'),
         starts_with('af_median'))
```


```{r}
summary_01_04 |>
  mutate(imputed = (is.na(R_01) | is.na(R_04))) |>
  replace_na(list(R_01 = 0, R_04 = 0)) |>
  ggplot(aes(x = R_01, y = R_04)) +
  geom_point(aes(fill = imputed), alpha = 0.5, shape = 21) +
  scale_fill_manual(values = c('FALSE' = 'grey40', 'TRUE' = 'firebrick')) +
  theme_bw() +
  labs(x = "Intra-twin Pearson's R in tailored Hg38+rDNA",
       y = "Intra-twin Pearson's R in Paralkar's T2T+rDNA") +
  theme(
    aspect.ratio = 1,
    legend.position = 'none')
```

```{r}
summary_01_04 |>
  mutate(imputed = (is.na(af_median_01) | is.na(af_median_04))) |>
  replace_na(list(af_median_01 = 0, af_median_04 = 0)) |>
  ggplot(aes(x = af_median_01, y = af_median_04)) +
  geom_point(aes(fill = imputed), alpha = 0.5, shape = 21) +
  scale_fill_manual(values = c('FALSE' = 'grey40', 'TRUE' = 'firebrick')) +
  theme_bw() +
  labs(x = "Median Variant Frequency in tailored Hg38+rDNA",
       y = "Median Variant Frequency in Paralkar's T2T+rDNA") +
  theme(
    aspect.ratio = 1,
    legend.position = 'none')
```

### 03 vs 04


```{r}
summary_03_04 <- full_join(
  summary_03,
  summary_04,
  by = c('pos_ref', 'ref', 'alt'), suffix = c('_03', '_04')) |>
  arrange(pos_ref, ref, alt) |>
  select(pos_ref, ref, alt, 
         starts_with('num_pairs'), 
         starts_with('R_'),
         starts_with('af_median'))
```


```{r}
summary_03_04 |>
  mutate(imputed = (is.na(R_03) | is.na(R_04))) |>
  replace_na(list(R_03 = 0, R_04 = 0)) |>
  ggplot(aes(x = R_03, y = R_04)) +
  geom_point(aes(fill = imputed), alpha = 0.5, shape = 21) +
  scale_fill_manual(values = c('FALSE' = 'grey40', 'TRUE' = 'firebrick')) +
  theme_bw() +
  labs(x = "Intra-twin Pearson's R in Paralkar's Hg38+rDNA",
       y = "Intra-twin Pearson's R in Paralkar's T2T+rDNA") +
  theme(
    aspect.ratio = 1,
    legend.position = 'none')
```



```{r}
summary_03_04 |>
  mutate(imputed = (is.na(af_median_03) | is.na(af_median_04))) |>
  replace_na(list(af_median_03 = 0, af_median_04 = 0)) |>
  ggplot(aes(x = af_median_03, y = af_median_04)) +
  geom_point(aes(fill = imputed), alpha = 0.5, shape = 21) +
  scale_fill_manual(values = c('FALSE' = 'grey40', 'TRUE' = 'firebrick')) +
  theme_bw() +
  labs(x = "Median Variant Frequency in Paralkar's Hg38+rDNA",
       y = "Median Variant Frequency in Paralkar's T2T+rDNA") +
  theme(
    aspect.ratio = 1,
    legend.position = 'none')
```


